---
title: "Olympics_EDA_Presentation "
output: 
  flexdashboard::flex_dashboard:
    orientation: row
    vertical_layout: scroll
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(flexdashboard)
#loading reqiuerd packages
library(plyr)
library(dplyr)
library(tidyverse)
library(skimr)
library(rworldmap)
library(repr)
library(DT)
library(janitor)
library(GGally)
library(here)
library(kableExtra)
library(tidytuesdayR)

#import data 
tuesdata <- tidytuesdayR::tt_load("2021-07-27")
# Individual dataframes
olympics <- tuesdata$olympics

# A second dataframe is available
# It is not required, but may prove useful in certain cases
regions <- tuesdata$regions
olympics$medal <- factor(olympics$medal,levels = c("Gold","Silver","Bronze"))
olympics$season <- factor(olympics$season,levels = c("Summer","Winter"))
olympics$sex <-  factor(olympics$sex,levels = c("M","F"))
```

# Olympics

<br> **EDA analysis of the Olypics games** <br> *By* <br> Mohammed
Darwish, Salman Al-Harbi, Amal alghamdi, Abdulmalik Alrebdi

**Content**

1.  DATA
2.  EDA
3.  Problems and challenges
4.  Key takeaways

# DATA

### Short description of the data

This is a historical dataset on the modern Olympic Games, including all
the Games from Athens 1896 to Rio 2016, featuring summer and winter
sports competitions in which thousands of athletes from around the world
participate in a variety of competitions dataset contains 271116 rows
and 15 columns. The Olympics is more than just a quadrennial multi-sport
world championship. It is a lens through which to understand global
history. goals is to shed light on major patterns in Olympic history.
How many athletes, sports, and nations are there? Where do most athletes
come from? Who wins medals?

**The dataset contains the following variables:**

| Variable | Type      | Description                                   |
|----------|-----------|-----------------------------------------------|
| id       | double    | Athlete unique ID                             |
| name     | character | Athlete Name                                  |
| sex      | character | Athlete Sex                                   |
| age      | double    | Athlete Age                                   |
| height   | double    | Athlete Height (cm)                           |
| weight   | double    | Athlete weight (kg)                           |
| team     | character | Team name, typically a country name           |
| noc      | character | NOC 3-letter country code                     |
| games    | character | 4-digit year + season                         |
| year     | double    | 4-digit year                                  |
| season   | character | Season                                        |
| city     | character | Host city                                     |
| sport    | character | Sport                                         |
| event    | character | Sport + Event Name                            |
| medal    | character | Medal: Gold, Silver, Bronze, (plus NA values) |

**You can access the dataset directly from base package.**

## Row 1 {.tabset data-height="1000,"}

### **The detailed summary for every type as follow**

```{r}
olympics %>% skimr::skim() 

```

# EDA

## Row {data-height="350"}

### Findings

The US has the highest number of medals this may be a result of the high
population, the diversity, which enables them to have more options in
each sport and have a high number of athletes as we can see in 2.1.6

### **Top 20 teams with respect to number of medals**

```{r}
medal_counts <- olympics %>% filter(!is.na(medal))%>%
  group_by(team, medal) %>%
  summarize(Count=length(medal))

# order team by total medal count
levs_art <- medal_counts %>%
  group_by(team) %>%
  summarize(Total=sum(Count)) %>%
  arrange(Total) %>%
  select(team) %>% 
  tail(n=20)

medal_counts$team <- factor(medal_counts$team, levels=levs_art$team)
medal_counts <- medal_counts %>% filter(!is.na(team))


# plot
ggplot(medal_counts, aes(x=team, y=Count, fill=medal)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values=c("gold1","gray70","gold4")) +
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Team",y="Number of Medals athletes since 1896")
```

## Row

### Number of athletes per countery

```{r}
olympics %>%
  group_by(noc) %>%
  count(name) %>%
  summarise(count = sum(n)) %>% 
  arrange(count) %>% 
  tail(n=20) %>% 
  
#plot
ggplot(aes(x= reorder(noc,count), y=count)) +
 geom_col() +
 coord_flip() +
 labs(x="Team",y="Number of Olympics athletes since 1896")

```

### who got the most medals

```{r}
olympics %>% 
  filter(!is.na(medal)) %>% 
  group_by(name) %>% 
  summarise(name , total = length(medal)) %>%  distinct() %>% 
  arrange( desc(total)  ) %>%  head(10)  -> medalyPerson


medalyPerson %>% 
ggplot(aes(reorder(name,total) , total)) + geom_bar(stat =  "identity") +theme(axis.text.x = element_text(angle = 90)) + coord_flip()
```

## Row 3 {data-height="350"}

### Findings

Interest NOCs are leading the way for gender equality in the Olympics,so
some NOCs even sent a majority of female athletes to the Games ,In 1996
and 2016. The chart shows asimprovement in female participation from
1956 to 2016 .

### know lets take a look if number of female athlete has increased since 1897 and if did why ?

```{r}
#count the number of females and males for each year
counts_sex <- olympics %>% 
  group_by(year, sex) %>%
  summarize(Count = length(unique(id))) %>%
  spread(sex, Count)

#clean the data
counts_sex$M[is.na(counts_sex$M)] <- 0
counts_sex$F[is.na(counts_sex$F)] <- 0
counts_sex$year <- as.factor(counts_sex$year)

#calculate the total on participants 
counts_sex <- mutate(counts_sex,Total = sum(M,F,na.rm=T)) 

#calculate the percentage of females and males from the total
counts_sex$M <- counts_sex$M/counts_sex$Total
counts_sex$F <- counts_sex$F/counts_sex$Total

#creating the graph
names(counts_sex)[2:3] <- c("Male","Female")
counts_sex <- gather(counts_sex,key= sex , value = Count,2:3)

ggplot(counts_sex,aes(x=year,y=Count,fill=sex)) +
  geom_col() +
  coord_flip() 
```

## Row {data-height="150"}

<br> <br> <br> <br> <br>

### Findings

**changes of athletes' height,weight and BMI over time** height, weight
has a similar trend it started increase from 1965 and before that the
plot shows that the athlete tend to have lower weights, they might
thought that decrease in the weight would be better . one thing to
notice is that before 1950 there were lots of events that did not need
athlete such as art which might affect the result .\

<br> <br> <br> <br> <br> <br> <br> <br>

## changes of athletes' height,weight and BMI over time {.tabset}

### Height

```{r}
olympics %>% 
  filter(!is.na(height)) %>% 
  group_by(year , sex) %>% 
  summarise( height= mean(height))-> avgHight

ggplot(avgHight , aes(year , height , color = sex)) + geom_point()+ geom_smooth(se = FALSE)
```

### Weight

```{r}
olympics %>% 
  filter(!is.na(weight)) %>% 
  group_by(year , sex) %>% 
  summarise( weight= mean(weight))-> avgWeight

ggplot(avgWeight , aes(year , weight , color= sex)) + geom_point()+ geom_smooth(se = FALSE)
```

### BMI

```{r}
avgBMI <- avgHight
avgBMI$weight <- avgWeight$weight

avgBMI <- avgBMI %>% 
  mutate(bmi = (weight)/(height/100 * height/100))

ggplot(avgBMI ,aes(year , bmi , color = sex) )+ geom_point()+ geom_smooth(se = FALSE)
```

### Example

```{r}
olympics %>%
  filter(grepl('Art', sport)) ->artAge
artAge %>%
  group_by(year) %>%
  summarise(year, age) %>%
 filter(!is.na(age)) ->artAge

ggplot(artAge, aes(year, age)) +
  geom_jitter(width = 0.2, alpha = 0.1, shape = 16) +
  stat_summary(fun.data = mean_sdl,
               fun.args = list(mult = 1))
```

## Row 3 {data-height="350"}

### Findings

this shows that the number of Games decrease in the beginning of WWII
and then it increases dynamically after almost 5 years

### **Did WW2 affect the number of participated countries ?**

```{r, echo=FALSE}
medal2 <-
  olympics %>%
  group_by(year, medal = medal=="Gold" , season) %>% #the idea here is when we group by medal = gold this will show us the number of games (one medal per game)
  summarise(
    TotalMedal = length(medal) ## i dont know whay na.rm do yet but when i remove it the graph change a lot
  )
medal2 %>%
  filter(medal , medal == TRUE)    ->medal ##the idea here is will count the number of true(medal == gold) in each year summer or
ggplot(medal , aes(year , TotalMedal  , color = season)) + geom_line() +
  geom_point(aes(x=1945, y=300), size=30, shape=1, color="gold4") +
  labs(title ="WWII & Olympcs " ,y = "Total Games", x = "YEAR of the event" , caption = "*the circles shows the start and the end of the war \n   ") +
  # please change the title to good name
  theme_bw() + theme(legend.position = c(.09,.85) , )
```

## Row 3 {data-height="350"}

### Findings

this shows the disperse of any German participant between 1939 and 1950
because of the war . The number of participation did not return to
before the war until 1967

### **Germany in wrold war 2**

```{r}
olympics %>%
  group_by(team , year ) %>%
  filter(grepl("Germany", team)) %>%
  # filter(team == "Germany" | team == "*Germany" ) %>%
  summarise( team , year , numberOfParticipants= length(name) ,season) %>%
  distinct() -> GermanyDataset
GermanyDataset  %>%
  ggplot(aes(year ,numberOfParticipants , fill = team )) + geom_bar(stat = "identity")+ ##  I dont like this plot but
  facet_grid(. ~ season) +
  labs(title = "Germany during WWII era" , caption = "between  1939 to 1950 germany did not play any olympics  ") +scale_color_brewer(palette = "Dark3")

```

## Row 3 {data-height="350"}

### Findings

xxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

### **Does the age of the athlete effect on his/her performance?**

*The avg age in each year*

```{r}
olympics %>% 
  select(year , team,age) %>%
  filter(!is.na(age)) %>% 
  group_by(year) %>% 
  summarise(age = mean(age)) -> avrgAgeInEachYear 

ggplot(avrgAgeInEachYear , aes(year , age)) + geom_point()+ geom_smooth(se = FALSE)

winnersVslosers <- olympics %>% 
  mutate(winner = if_else(is.na(medal),"loser","winner"))
```

### Findings

xxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

### How old are most of the winners and losers?

```{r}
winnersVslosers$winner <- factor(winnersVslosers$winner,levels = c("loser","winner")) 

winnersVslosers %>%
  filter(!is.na(age)) %>% 
  group_by(winner,age) %>% 
  summarise(count=n()) -> winnersVslosers

#plot
winnersVslosers %>% 
  ggplot(aes(x=age ,fill= winner)) +
  geom_density(alpha=0.25)
```

## Row 3

### Findings

xxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

## Row

### Mean of age in year With respect of kind of model

```{r}
medal <-
  olympics %>% 
  filter(!is.na(age),!is.na(medal)) %>% 
  group_by(year, medal) %>% 
  summarise(age = mean(age))

#plot
ggplot(medal , aes(x=year , y=age, color = medal)) + geom_line() +
scale_color_manual(values=c("gold1","gray70","gold4")) 
```

## Row 3 {data-height="150"}

### Findings

xxxxxxxxxxxxxxxxxx
xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxXXXXXXXXXXXXXXxxxxxxxxxxxxxxxxxxxxx
xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxx

## Row 3 {data-height="650"}

### kids who got Olympics models(less than 13)

```{r}
olympics %>% 
  filter(age<13, !is.na(medal)) %>% 
  select(-year,-season) %>% 
  arrange(desc(age)) %>%
  head(5) %>%
  knitr::kable()

```

## Row 3 {data-height="150"}

### Findings

xxxxxxxxxxxxxxxxxx
xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxXXXXXXXXXXXXXXxxxxxxxxxxxxxxxxxxxxx
xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxx

## Row 3 {data-height="1000"}

### kids who got Olympics models(less than 13)

```{r}
olympics %>%
  filter(age>65, !is.na(medal)) %>%
  select(-year,-season) %>% 
  arrange(desc(age)) %>%
  head(5) %>%
  knitr::kable() 
```

# Problems and challenges

***Winter and Summer Games were held in the same year up until 1992.
After that, they staggered them such that Winter Games occur on a four
year cycle starting with 1994, then Summer in 1996, then Winter in 1998,
and so on.*** <br> ***How to decide which plots should be presented***
<br> **There are only a few events that happened every year so it is
hard to evaluate the attributes of the athlete without deep analysis**
<br> **Using git we had some issues in pushing and pulling the content
and how we organize our work** <br> **In the early years, there is a lot
of missing data**

# Key Takeaways from the dataset
***The US has the highest number of medals and athletes.***<br>
***Brazil has the highest number of medals in football.***<br>
***Female participation in the Olympic Games is on the rise.***<br>
***The size of Olympians have become more extreme over time. ***<br>
***In most sports this means taller and heavier, but in a few sports such as gymnastics, athletes have become smaller.***<br>
***The proportion of winners athletes with ages below 60 is higher than that above 60.***<br>
***The Olympic Games plays a vital role in fostering unity among nations.***<br>
