---
title: "Olympics_EDA_Presentation "
output: 
  flexdashboard::flex_dashboard:
    orientation: row
    vertical_layout: scroll
    theme: united
    code_folding: hide


---

```{r setup, include=FALSE}
library(flexdashboard)
#loading reqiuerd packages
library(plyr)
library(dplyr)
library(tidyverse)
library(skimr)
library(rworldmap)
library(repr)
library(DT)
library(janitor)
library(GGally)
library(here)
library(kableExtra)
library(tidytuesdayR)

#import data 
tuesdata <- tidytuesdayR::tt_load("2021-07-27")
# Individual dataframes
olympics <- tuesdata$olympics

# A second dataframe is available
# It is not required, but may prove useful in certain cases
regions <- tuesdata$regions
olympics$medal <- factor(olympics$medal,levels = c("Gold","Silver","Bronze"))
olympics$season <- factor(olympics$season,levels = c("Summer","Winter"))
olympics$sex <-  factor(olympics$sex,levels = c("M","F"))
```

# DATA 
### Short description of the data
aaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaa
aaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaa
aaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaa
aaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaa

**The dataset contains the following variables:**


| Variable | Type      | Description                                    |
|----------|-----------|------------------------------------------------|
| id       | double    | Athlete unique ID                              |
| name     | character | Athlete Name                                   |
| sex      | character | Athlete Sex                                    | 
| age      | double    | Athlete Age                                    |
| height   | double    | Athlete Height (cm)                            | 
| weight   | double    | Athlete weight (kg)                            |
| team     | character | Team name, typically a country name            |
| noc      | character | NOC 3-letter country code                      |
| games    | character | 4-digit year + season                          |
| year     | double    | 4-digit year                                   | 
| season   | character | Season                                         |
| city     | character | Host city                                      | 
| sport    | character | Sport                                          |
| event    | character | Sport + Event Name                             |
| medal    | character | Medal: Gold, Silver, Bronze, (plus NA values)  |


**You can access the dataset directly from base package.**

## Row 1 {data-height= 650}

**The detailed summary for every type as follow**


```{r}
DT::datatable(olympics, 
              rownames = FALSE, 
              options = list(pageLength = 5, scrollX = TRUE), 
              class = 'white-space: nowrap' )
olympics %>% skim_without_charts()
```


# EDA
How did the numbers of female and male athlete changed over time


### Plot 1

```{r, echo=FALSE}
print("Hello")
```


### Plot 2

**changes of athletes' height,weight and BMI over time**

## changes of athletes' height,weight and BMI over time {.tabset}

### Height
```{r}
olympics %>% 
  filter(!is.na(height)) %>% 
  group_by(year) %>% 
  summarise( height= mean(height))-> avgHight

ggplot(avgHight , aes(year , height)) + geom_point()+ geom_smooth(se = FALSE)
```

### Weight
```{r}
olympics %>% 
  filter(!is.na(weight)) %>% 
  group_by(year) %>% 
  summarise( weight= mean(weight))-> avgWeight

ggplot(avgWeight , aes(year , weight)) + geom_point()+ geom_smooth(se = FALSE)
```

### BMI

```{r}
avgBMI <- avgHight
avgBMI$weight <- avgWeight$weight

avgBMI <- avgBMI %>% 
  mutate(bmi = (weight)/(height/100 * height/100))

ggplot(avgBMI ,aes(year , bmi) )+ geom_point()+ geom_smooth(se = FALSE)
```



## Row 2

### did WW1 and WW2 affect the number of participated countries ?

**How did the Olympics suffer from the WWII**

 
```{r, echo=FALSE}
medal2 <-
  olympics %>% 
  group_by(year, medal = medal=="Gold" , season) %>% #the idea here is when we group by medal = gold this will show us the number of games (one medal per game)
  
  summarise(
    TotalMedal = length(medal) ## i dont know whay na.rm do yet but when i remove it the graph change a lot
  )


medal2 %>% 
  filter(medal , medal == TRUE)    ->medal ##the idea here is will count the number of true(medal == gold) in each year summer or 

ggplot(medal , aes(year , TotalMedal  , color = season)) + geom_line() +
  geom_point(aes(x=1920, y=470), size=30, shape=1, color="gold4") +
  geom_point(aes(x=1945, y=300), size=30, shape=1, color="gold4") +
  labs(title ="WWII & Olympcs " ,y = "Total Games", x = "YEAR of the event" , caption = "*the circles shows the start and the end of the war \n this also shows that the winter olympics started in the middle of the war  ") +
  # please change the title to good name
  theme_bw() + theme(legend.position = c(.05,.92) , )
```

## Row

### Germany in wrold war 2 




```{r}
olympics %>% 
  group_by(team , year ) %>% 
  filter(grepl("Germany", team)) %>% 
  
  # filter(team == "Germany" | team == "*Germany" ) %>% 
  summarise( team , year , numberOfParticipants= length(name) ,season) %>% 
  distinct() -> GermanyDataset

GermanyDataset  %>% 
  ggplot(aes(year ,numberOfParticipants , color = team )) + geom_bar(stat = "identity")+ ##  I dont like this plot but  
  
  facet_grid(. ~ season) +
  labs(title = "Germany during WWII era" , caption = " between 1920 to 1945 Germany have not played any olympics except 3   ") 

```

## Row 2

### **Does the age of the athlete effect on his/her performance?**
*The avg age in each year*

```{r}
olympics %>% 
  select(year , team,age) %>%
  filter(!is.na(age)) %>% 
  group_by(year) %>% 
  summarise(age = mean(age)) -> avrgAgeInEachYear 

ggplot(avrgAgeInEachYear , aes(year , age)) + geom_point()+ geom_smooth(se = FALSE)

winnersVslosers <- olympics %>% 
  mutate(winner = if_else(is.na(medal),"loser","winner"))
```

### **How old are most of the winners and losers?**

```{r}
winnersVslosers$winner <- factor(winnersVslosers$winner,levels = c("loser","winner")) 

winnersVslosers %>%
  filter(!is.na(age)) %>% 
  group_by(winner,age) %>% 
  summarise(count=n()) -> winnersVslosers

#plot
winnersVslosers %>% 
  ggplot(aes(x=age ,fill= winner)) +
  geom_density(alpha=0.25)
```

## Row 
### Top 10 teams with respect to number of medals
```{r}
# count number of medals for each team
medal_counts <- olympics %>% filter(!is.na(medal))%>%
  group_by(team, medal) %>%
  summarize(Count=length(medal))

# order team by total medal count
levs_art <- medal_counts %>%
  group_by(team) %>%
  summarize(Total=sum(Count)) %>%
  arrange(Total) %>%
  select(team) %>% 
  tail(n=20)

medal_counts$team <- factor(medal_counts$team, levels=levs_art$team)
medal_counts <- medal_counts %>% filter(!is.na(team))


# plot
ggplot(medal_counts, aes(x=team, y=Count, fill=medal)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values=c("gold1","gray70","gold4")) +
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Team",y="Number of Medals athletes since 1896")
```

### With respect with medal type *(gold , silver ,bronze)*
```{r}
medal_countsALL <- olympics %>% filter(!is.na(medal))%>%
  group_by(team) %>%
  summarize(Count = length(medal)) 

medal_countsALL <- medal_countsALL[order(medal_countsALL$Count, decreasing = TRUE), ] %>% head(10)

ggplot(medal_countsALL , aes(Count , team )) + geom_bar(stat =  "identity")
```

## Row 

### **Number of athletes per countery**
```{r}
olympics %>%
  group_by(noc) %>%
  count(name) %>%
  summarise(count = sum(n)) %>% 
  arrange(count) %>% 
  tail(n=20) %>% 
  
#plot
ggplot(aes(x= reorder(noc,count), y=count)) +
 geom_col() +
 coord_flip() +
 labs(x="Team",y="Number of Olympics athletes since 1896")

```

## Row
### Mean of age in year With respect of kind of model 

```{r}
medal <-
  olympics %>% 
  filter(!is.na(age),!is.na(medal)) %>% 
  group_by(year, medal) %>% 
  summarise(age = mean(age))

#plot
ggplot(medal , aes(x=year , y=age, color = medal)) + geom_line() +
scale_color_manual(values=c("gold1","gray70","gold4")) 
  



```


# Problems and challenges you’ve encountered
**blank**

***Winter and Summer Games were held in the same year up until 1992. After that, they staggered them such that Winter Games occur on a four year cycle starting with 1994, then Summer in 1996, then Winter in 1998, and so on.***

***How to decide which plots should be presented***

# Key Takeaways from the dataset backed by visualizations and/or tables
